Interactive Filter and Display of Hillary Clinton's Emails: \\A Cautionary Tale of Metadata

### 2017-03-25 VAST_submission/vgtc_journal_latex/paper.tex
More paper updates

In addition to the spoked network graph, a time series plot of daily email volume is displayed. This plot has two lines, the first is a line showing the total volume of emails for each day, and the second shows the number of emails satisfying the classification filter criteria for a given day. This allows users to view which days have an exceptionally high or low number of classified conversations. Returning to the motivation of hypothesis generation, referencing world events on any days of interest provides a number of interesting insights on how public figures like Clinton react and respond to world events.

@@

>>>>

In addition to the spoked network graph, a time series plot of daily email volume is displayed. This plot has two lines, the first is a line showing the total volume of emails for each day, and the second shows the number of emails satisfying the classification filter criteria for a given day. This allows users to view which days have an exceptionally high or low number of classified conversations. Returning to the motivation of hypothesis generation, referencing world events on any days of interest provides a number of interesting insights on how public figures like Clinton react and respond to world events. To augment this graph, Clinton's official foreign visit schedule (https://history.state.gov/departmenthistory/travels/secretary/clinton-hillary-rodham) was extracted, cleaned, and incorporated into the application. This coloured time segments during which Clinton was travelling with blue rectangles, highlighting her pattern of foreign visits and allowing a comparison between these and the daily email sending patterns.

@@

\section{Observations}

In designing, creating, testing, and playing with the app, a number of interesting observations were made. These are summarized below, and some address contentious issues in regards to Clinton's emails and a number of significant events which occurred during her tenure as the United States Secretary of State.

\subsection{Daily Email Patterns}

\includegraphics{SentReceivedEntireTenure}

The most striking thing noticed about the email sending times was the regularity of the pattern. Across the entirety of Clinton's four year tenure as the United States Secretary of State, email communication is densest between midnight and four in the morning, and there is a conspicuous gap in communication between roughly four in the afternoon and ten at night. This pattern is quite surprising, as it suggests that Clinton and her staff are most actively communicating late at night, rather than at the start of the regular work day around seven or eight in the morning. It would be expected that individuals would be sleeping late at night, and as a result that the density of emails would be relatively low. Rather, the data suggests that the only real break Clinton and her staff take from email communication is in the early evening, when many of them doubtless have official functions to attend. Regardless of the individual daily variation, the consistency of this pattern suggests that Clinton and her staff slept at strange times, and perhaps very irregularly, or were at least frequently interrupted by emails while sleeping. It is also possible these late night emails are more common because they are the ones Clinton chose to release, after all, it has been observed Clinton used the service Bleachbit to delete many old emails from her server (http://www.politico.com/story/2016/08/hillary-clinton-emails-bleachbit-227425, http://money.cnn.com/2016/08/26/technology/hillary-clinton-bleachbit/) and the organizational emails sent late at night to coordinate the following day may have simply been overlooked or maintained deliberately. Either way, the presence of such high volumes late at night tells a great deal about the irregularity of sleep that must have occurred among Clinton's staff during her tenure.

Another interesting pattern in the data is that of the two alternating modes at two and three in the morning. Immediately, hypothesizing that the cause of these modes is some regular server function is obvious due to the precision and consistency of the pattern. The alternation in the pattern was initially vexing, but comparing the shifts in the mode to the dates of daylight savings time in North America for the time period found a perfect match. It would seem that the alternation, then, is not deliberate, but rather a consequence of the one hour time shift that occurs in March and November every year in North America. Still, this observation does not reveal much about the content of the emails sent at this time, or why they are so regular. To better explore these emails, the app was run on a new set of data which was filtered to only contain the emails sent at two and three in the morning. Of the 32,359 emails sent during Clinton's tenure in this data set, 4,020 or roughly one eighth were sent at exactly two or three in the morning.
